---
title: "p8105_hw2_zq2209"
author: "Zining Qi"
date: "2022-10-03"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
library(tidyr)
```

# Problem 1
```{r}
trans_ent = 
  read_csv(
    "/Users/qizining/Desktop/p8105_hw2_zq2209/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route",
    values_to = "route_value"
  ) %>%
  filter(!is.na(route_value)) %>%
  unique() %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The table is 1582 x 20. There are 11 routes variable and 9 other variables, which are line, station, name, station latitude / longitude, served, entry, vending, entrance type, and ADA compliance. So far, the cleaning steps are clean names, select required variables and change entry to logic variable. 


```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

There are 465 rows in this table, which means there are 465 unique stations.

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 84 rows in this table, which means there are 84 stations with ada compliant.

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

There are 33.10% without vending entrance.

```{r}
data1 = read_csv(
    "/Users/qizining/Desktop/p8105_hw2_zq2209/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>%  
  pivot_longer(
    route1:route11,
    names_to = "route",
    values_to = "route_value") %>% 
  filter(route_value == "A") %>% 
  select(station_name, line) %>% 
  distinct
```

There are 60 rows.

```{r}
data2 = read_csv(
    "/Users/qizining/Desktop/p8105_hw2_zq2209/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route",
    values_to = "route_value") %>% 
  filter(route_value == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 17 rows



# Problem 2
```{r}
## read and clean Mr Trash Wheel sheet
mr_trash_wheel_raw = read_excel("Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N550", col_names = TRUE)
mr_trash_wheel = mr_trash_wheel_raw %>% drop_na() %>%
  janitor::clean_names() %>%
  mutate(sports_balls = as.integer(round(sports_balls, 0)),
         dumpster = as.numeric(dumpster),
         homes_powered = as.integer(round(homes_powered, 0)),
         year = as.numeric(year),
         name = "Mr. Trash Wheel") %>%
  select(name, everything())
```

```{r}
## read and clean Professor Trash Wheel sheet
prof_trash_wheel_raw = read_excel("Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M97", col_names = TRUE)
prof_trash_wheel = prof_trash_wheel_raw %>% drop_na() %>%
  janitor::clean_names() %>%
  mutate(dumpster = as.numeric(dumpster),
         homes_powered = as.integer(round(homes_powered, 0)),
         name = "Professor Trash Wheel") %>%
  select(name, everything())
```



```{r}
# Merging two data sets
merged_trash_wheel = bind_rows(mr_trash_wheel, prof_trash_wheel) %>%
  janitor::clean_names()
nrow(merged_trash_wheel)
ncol(merged_trash_wheel)
```

There are 568 rows and 15 columns in merged dataset. The variables are name, dumpster, month, year, date, weight_tons, "volume_cubic_yards" "plastic_bottles" "polystyrene", "cigarette_butts" "glass_bottles" "grocery_bags"  "chip_bags" "sports_balls" "homes_powered"    
      
```{r}
# Total weight of trash collected by Professor Trash Wheel
total_number_of_weight = merged_trash_wheel %>% 
  filter(name == "Professor Trash Wheel") %>%
  select(weight_tons) %>%
  colSums()
total_number_of_weight
```

The number of weight by professor trash wheel is 162.54.

```{r}
# Total number of sports balls collected by Mr. Trash Wheel in 2020
total_sportball_mr_2022 = mr_trash_wheel[mr_trash_wheel$year == "2020", ] %>%
  select(sports_balls) %>%
  colSums()
total_sportball_mr_2022
```

The number of sports ball by Mr. Trash wheel is 856.



# Problem 3
```{r}
# Data cleaning in pols-month.csv
pols = read_csv("/Users/qizining/Desktop/p8105_hw2_zq2209/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = recode(month, "01" = "Janurary", "02" = "Feburary", "03" = "March", "04" = "April", "05" = "May", "06" = "June", "07" = "July", "08" = "August", "09" = "September", "10" = "October", "11" = "November", "12" = "December")) %>%
  pivot_longer(c(prez_gop, prez_dem), names_to = "president", values_to = "value") %>%
  mutate(value = replace(value, value == 0, NA ),
         president = recode(president, "prez_gop" = "gop", "prez_dem" = "dem")) %>%
  drop_na() %>%
  select(-c(day, value))

nrow(pols)
ncol(pols)
```
 
There are 822 rows and 9 columns in this cleaned dataset. The dataset range from 1947 to 2015. And the variables are "year" "month" "gov_gop" "sen_gop" "rep_gop" "gov_dem" "sen_dem"   "rep_dem"   "president". 
 
```{r}
# Data cleaning in snp.csv
snp = read_csv("/Users/qizining/Desktop/p8105_hw2_zq2209/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(month = recode(month, "1" = "Janurary", "2" = "Feburary", "3" = "March", "4" = "April", "5" = "May", "6" = "June", "7" = "July", "8" = "August", "9" = "September", "10" = "October", "11" = "November", "12" = "December"),
         year = as.numeric(year),
         year = ifelse(year >= 50, year + 1900, year + 2000),
         ) %>%
  mutate(year = as.character(year)) %>%
  arrange(year, month) %>%
  select(year, month, close)

nrow(snp)
ncol(snp)
```

There are 787 rows and 3 columns in this cleaned dataset. The dataset range from 1950 to 2015. And the variables are "year" "month" "close".

```{r}
# Data cleaning in unemployment.csv
unemployment = read_csv("/Users/qizining/Desktop/p8105_hw2_zq2209/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  rename(Janurary = jan, Feburary = feb, March = mar, April = apr, May = may, June = jun, July = jun, August = aug, September = sep, October = oct, November = nov, December = dec) %>%
  pivot_longer(
    Janurary:December,
    names_to = "month",
    values_to = "unemployment"
  ) %>%
  mutate(year = as.character(year)) %>%
  arrange(year, month)

nrow(unemployment)
ncol(unemployment)
```

There are 816 rows and 3 columns in this cleaned dataset. The dataset range from 1948 to 2015. And the variables are "year" "month" "unemplyment".

```{r}
merged_data = 
  pols %>% full_join(snp, by = c("year", "month"))
```
```{r}
full_data =
  merged_data %>% full_join(unemployment, by = c("year", "month"))

nrow(full_data)
ncol(full_data)
```

There are 896 rows and 11 columns in this merged dataset. The dataset range from 1947 to 2015. And the variables are "year" "month" "unemplyment" "gov_gop" "sen_gop" "rep_gop" "gov_dem" "sen_dem"   "rep_dem"   "president" "close".











